---
title: "Model update"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(Hmisc)
library(rms)
library(glmnet)
library(tidyr)
library(dplyr)
library(purrr)
library(ROCit)
```

This notebook updates iteratively updates the base prognostic model applying recalibration, by adding additional lab variables in blocks of 2 days

Some ideas to consider further:
* If Elastic Net variable selection is applied on data w/out missing data,
  generate frames by censoring a given percentage of data (e.g.: 10%) and apply MI
* Try Bayesian Variable Selection:
https://github.com/BasisResearch/millipede
https://cran.r-project.org/web/packages/BayesVarSel/
https://cran.r-project.org/web/packages/BayesVarSel/BayesVarSel.pdf

```{r message=FALSE, cache=F}
source('./mdl-evaluation.R')
source('./mdl-selection.R')

num_imp <-100       # Number of multiple imputations
num_boot <- 200     # Number of bootstrap samples
random_seed <- 161  # Setting random seed for reproducibility

min_age <- 18       #  Minimum age for inclusion in analysis

sel_thres <- 0.7    # Minimum fraction of models in which a variable appears as a positive predictor in order to be selected

max_cor <- 0.65     # Maximum correlation between variables to be included in model
```

```{r}
# Load the data, remove children, remove patients with less than 1 day of stay, and transform Ct values into Z-scores

data <- read.csv('./data/data-cleaned.csv')
data <- data %>% filter(age >= min_age)
data <- data %>% filter(days > 0)

data$ct12 <- (data$ct12 - mean(data$ct12, na.rm=T))/sd(data$ct12, na.rm=T)
data$ct34 <- (data$ct34 - mean(data$ct34, na.rm=T))/sd(data$ct34, na.rm=T)
data$ct56 <- (data$ct56 - mean(data$ct56, na.rm=T))/sd(data$ct56, na.rm=T)
data$ct78 <- (data$ct78 - mean(data$ct78, na.rm=T))/sd(data$ct78, na.rm=T)

data
```

```{r}
# Construct the base prognostic model that only uses age and Ct in the first 48 hours

base_model <- function(df) {
  # Calculate the linear term in the logistic function of the model
  # p(x) = 1 / (1+exp(-lp(x))
  lp = with(df, 
            0.86215 - 
              0.05343*age + 
              3*10^-5*max(age - 5.0, 0)^3 - 
              6*10^-5*max(age - 30.0, 0)^3 + 
              3*10^-5*max(age - 58.0, 0)^3 - 
              0.8694*ct12)
  return(lp)
}
```

```{r}
# Add the logit contribution from the base model to the dataframe

vars0 <- c('out', 'age', 'ct12')
all_data0 <- data %>% select(all_of(vars0)) # This dataframe includes missing entries

lp0 <- unlist(map(1:nrow(all_data0), function(x) base_model(all_data0[x,])))
data['lp0'] <- lp0
data
```


```{r}
# Plot calibration and ROC plots of the base model

mdl_folder0 <- "./models/model-0"
dir.create(mdl_folder0)

data0 <- all_data0[complete.cases(all_data0),]

pred0 <- unlist(map(1:nrow(data0), function(x) base_model(data0[x,]))) # prediction in log-odds (probabilities on logit scale)
prob0 <- 1 / (1 + exp(-pred0))                                         # prediction as probability response

# Generate the data frame with the data needed for the calibration and ROC plots
plot_data0 <- data.frame(out = data0$out, logodds = pred0, prob = prob0)

cal_plot(base_model, plot_data0, "prob", model_label = '', dir = mdl_folder0, fn = "cal_plot.pdf")
cal_c0 <- cal_coeff(plot_data0, "prob", mdl_folder0, "cal_plot_coefficients.txt")
print(cal_c0)

roc_plot(data0$out, prob0, dir = mdl_folder0, fn = "roc_plot.pdf")
```

```{r}
plot_roc_perf(plot_data0$out, plot_data0$prob)
```

```{r}
conf_matrix(plot_data0$out, plot_data0$prob, 0.5, mdl_folder0, "conf_matrix.txt")
```

==============================================================================================
UPDATED MODEL FOR DAYS 3-4

```{r}
# Train the model for days 3-4
mdl_vars34 <- c('lp12', 'ct34')

mdl_imp_formula34 <- "~out + lp0 + ct34"
mdl_lgr_formula34 <- paste('out', paste(mdl_vars34, collapse=" + "), sep=" ~ ")
mdl_lgr_formula34

mdl_folder34 <- "./models/model-days34"
dir.create(mdl_folder34)

model34 <- generateModel(random_seed, num_imp, num_boot, 3, mdl_imp_formula34, mdl_lgr_formula34, data34, mdl_vars34, mdl_folder34)

# Add the logit contribution from the model for days 3-4 to the dataframe
data['lp34'] <- predict(model34, data)
```

```{r}
# Generate the calibration and ROC plots for the models for days 3-4

vars <- append(mdl_vars34, 'out')
df <- data %>% select(all_of(vars))
full_data34 <- df[complete.cases(df),]
full_data34

pred34 <- predict(model34, full_data34) # prediction in log-odds (probabilities on logit scale)
probs34 <- 1 / (1 + exp(-pred34))  # prediction as probability response

# Generate the data frame with the data needed for the calibration and ROC plots
plot_data34 <- data.frame(out = full_data34$out, logodds = pred34, prob = probs34)

cal_plot(model34, plot_data34, "prob", model_label = '', dir = mdl_folder34, fn = "cal_plot.pdf")
cal_c34 <- cal_coeff(plot_data34, "prob", mdl_folder34, "cal_plot_coefficients.txt")
print(cal_c34)

roc_plot(plot_data34$out, plot_data34$prob, dir = mdl_folder34, fn = "roc_plot.pdf")
```

```{r}
plot_roc_perf(plot_data34$out, plot_data34$prob)
```

```{r}
conf_matrix(plot_data34$out, plot_data34$prob, 0.36, mdl_folder34, "conf_matrix.txt")
```

==============================================================================================
UPDATED MODEL FOR DAYS 5-6

```{r}
# Train the model for days 5-6 on the selected variables
mdl_vars56 <- c('lp34', 'ct56')

mdl_imp_formula56 <- "~out + lp34 + ct56"

mdl_lgr_formula56 <- paste('out', paste(mdl_vars56, collapse=" + "), sep=" ~ ")
mdl_lgr_formula56

mdl_folder56 <- "./models/model-days56"
dir.create(mdl_folder56)

# setting number of knots to zero based on this discussion:
# https://stats.stackexchange.com/questions/117383/impute-missing-values-using-aregimpute
model56 <- generateModel(random_seed, num_imp, num_boot, 0, mdl_imp_formula56, mdl_lgr_formula56, data56, mdl_vars56, mdl_folder56)

# Add the logit contribution from the model for days 5-6 to the dataframe
data['lp56'] <- predict(model56, data)
```

```{r}
# Generate the calibration and ROC plots for the models for days 5-6

vars <- append(mdl_vars56, 'out')
df <- data %>% select(all_of(vars))
full_data56 <- df[complete.cases(df),]
full_data56

pred56 <- predict(model56, full_data56) # prediction in log-odds (probabilities on logit scale)
probs56 <- 1 / (1 + exp(-pred56))  # prediction as probability response

# Generate the data frame with the data needed for the calibration and ROC plots
plot_data56 <- data.frame(out = full_data56$out, logodds = pred56, prob = probs56)

cal_plot(model56, plot_data56, "prob", model_label = '', dir = mdl_folder56, fn = "cal_plot.pdf")
cal_c56 <- cal_coeff(plot_data56, "prob", mdl_folder56, "cal_plot_coefficients.txt")
print(cal_c56)

roc_plot(plot_data56$out, plot_data56$prob, dir = mdl_folder56, fn = "roc_plot.pdf")
```

```{r}
plot_roc_perf(plot_data56$out, plot_data56$prob)
```

```{r}
conf_matrix(plot_data56$out, plot_data56$prob, 0.15, mdl_folder56, "conf_matrix.txt")
```

==============================================================================================
UPDATED MODEL FOR DAYS 7-8

DOES NOT WORK BECAUSE THE DATA IS HIGHLY CLASS-IMBALANCED (out of 39 cases, 37 survived are 2 died, leading to variable selection/model fitting to fail)